sensi_an_rotations201911. To analyse rotation runs for paper

### 2019-11-06 inst/documents/sensi_an_rotations201911.Rmd
starting plots for paper

---
title: "sensi_an_rotations1. Start of sensitivity analysis for rotations."
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

@@

  outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

10,000 runs comparing rotations and sequences.

```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#eval=FALSE just to temporarily hide this text

# Starting question : is there any benefit to rotations when there are no costs and no migration to an untreated area ?
# 
# Test 1. No refugia (i.e. Coverage=100%) or fitness costs. Would expect all policies to be equivalent.
# 
# Test 2. Dominance of resistance and fitness costs both set to 0.5.
# I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles. I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same. I’m probably wrong but it’s a nice test…..
#

#\pagebreak
```

@@

#start small & build it up
#maybe create an equiv. of resistance::sensiAnPaperPart.r later

# example for 1 run
# linputs <- list(max_gen = 50, n_insecticides = 2)
# do.call(run_rot, linputs)
# dfinputs <- as.data.frame(linputs)

# So one way of doing is to create an linputs list for each run & pass to run_rot()
# so run_rot() only ever runs for 1 scenario
# in contrast resistance::runModel2() accepted an input matrix & would run for multiple scenarios
# not sure which is better

# set random seed so that results can be reproduced
#set.seed(1)
set.seed(2)
max_gen         <- 500
plot_scenarios  <- FALSE
n_scenarios     <- 10 #000 
already_ran     <- FALSE
#already_ran     <- TRUE

# if already done runs, just load data for plots
if (already_ran)
{
  #C:/rsprojects/rotations/inst/documents/

  #load("df_res_all_rot10000.rda")
  load("df_res_all_rot_mig1000.rda")

}else
{
  df_res_all <- data.frame()

  for(i in 1:n_scenarios)
  {
    #message should output to the R markdown console
    if (i%%200 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())

    # inputs (could also be saved to a dataframe first & extracted one row at a time)
    linputs <- list()

    ## constant inputs ##
    linputs$coverage <- 1
    linputs$migration <- 0
    #linputs$cost <- 0 #variable cost below
    linputs$max_gen <- max_gen

    #beware min_rwr_interval, it's to stop repeated small rotations
    linputs$min_rwr_interval <- 5
    linputs$min_gens_switch_back <- 1
    linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
    linputs$no_r_below_mut <- FALSE #BEWARE trying out new mutation-selection balance option
    #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
    linputs$plot <-     FALSE

    ## variable inputs ##
    linputs$cost <-           runif(1, min=0,   max=0.1) 
    #linputs$cost <-           0 
    linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9
    linputs$male_expo_prop <- runif(1, min=0,   max=1)
    linputs$eff <-            runif(1, min=0.5, max=1) # min=0.3, max=1
    linputs$rr  <-            runif(1, min=0.1, max=0.9)
    linputs$dom_sel <-        runif(1, min=0,   max=1)
    linputs$dom_cos <-        runif(1, min=0,   max=1) 
    #TODO make this lognormal to get lower frequencies too
    linputs$start_freqs <-    runif(1, min=0.005,   max=0.1) #note not log yet   
    linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
    # dispersal (initially with cost set to 0)
    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
    #linputs$migration <-      runif(1, min=0.1, max=0.9)

    # create null objects used later for plotting
    ggrot <- ggseq <- NULL

    # to compare sequence to rotation run all the scenarios above for both seq & rot 
    for(rot_or_not in 0:1)
    {  
      #to get rot_int of 0 every other run
      if (rot_or_not == 0) linputs$rot_interval <- 0
      else   #random rot interval integer
          linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1

      ##################
      # run one scenario
      dfres <- do.call(run_rot, linputs)

      # summarise results per scenario
      # note this is also currently calc at end of run_rot() here is probably better

      res <- gens_under_thresh(dfres)

      # to plot results for checking
      if ( linputs$rot_interval == 0 )
      {
        gens_seq <- as.numeric(res)
      } else
      {
        gens_rot <- as.numeric(res)
      }

      if (plot_scenarios)
      {
        # make plot
        ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)

        # to plot results for checking
        if ( linputs$rot_interval == 0 )
        {
          ggseq <- ggres
          gens_seq <- as.numeric(res)
        } else
        {
          ggrot <- ggres
          gens_rot <- as.numeric(res)
        }
      }

    # TODO I may want to concat all dfres into one object (adda column for scenario num) to save for later analysis
    # e.g. if I wanted to calculate average resistance frequencies over time in addition to gens below threshold

    } #end rot_or_not loop

    # make an output row for each scenario with inputs, gens_rot, gens_seq
    # need to make sure I retain the rot_interval for rot

    # add scenario id column onto results
    df_in_out <- as.data.frame(c(linputs, gens_rot=gens_rot, gens_seq=gens_seq, id=i))
    df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq

    # previously I used rbind this scenario inputs & results onto previous
    # not a good speed or memory way of doing
    #df_res_all <- rbind(df_res_all, df_in_out)

    # now replaced with this better way
    # set a dataframe for the outputs to the required size in scenario1
    if (i==1)
    {
      df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=n_scenarios))
      names(df_res_all) <- names(df_in_out)
    }

    #putting results into one row of overall results dataframe
    df_res_all[i,] <- df_in_out

    if (plot_scenarios)
    {  
      # plot input vals
      # extract out those between 0 & 1
      dfi <- data.frame(t(as.data.frame(linputs)))
      names(dfi) <- 'value'
      dfi$inputs <- rownames(dfi)

      ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
                geom_point() +
                theme_minimal() +
                ylim(0,1) +
                theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))

      grid.arrange(ggrot, ggseq, ggins, ncol=1)

      cat('scenario',i,'sum generations deployed insecticide under thresh rotation:', gens_rot,' sequence:',gens_seq,"\n\n")

      #can I give the code to rerun this scenario
      cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
                  "cost=",signif(linputs$cost,2),", ",
                  "start_freqs=",signif(linputs$start_freqs,2),", ", 
                  "rot_interval=",linputs$rot_interval,", ",
                  "eff=",signif(linputs$eff,2),",\n",
                  "dom_sel=",signif(linputs$dom_sel,2),", ",
                  "dom_cos=",signif(linputs$dom_cos,2),", ",
                  "rr=",signif(linputs$rr,2),", ",
                  "expo_hi=",signif(linputs$expo_hi,2),", ",
                  "coverage=",signif(linputs$coverage,2),",\n",
                  "migration=",signif(linputs$migration,2),", ",
                  "max_gen=",linputs$max_gen,", ", 
                  "min_rwr_interval=",linputs$min_rwr_interval,            
                  ")\n") )

    #I could save partial outputs every 200 scenarios
    #just in case doesn't make to end
    if (i%%1000 == 0)
    {
      #df_res_all$rot_minus_seq <- df_res_all$gens_rot - df_res_all$gens_seq

      # save object containing inputs and results as rda for analysis
      save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
    }

  } #end scenario loop

@@

## mostly (~67%) no difference but infrequent where rotations better
## ~ 7% of runs, rotation at least 20% better than sequence

@@

```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=3.5}
#eval false because smaller runs don't have these

@@

#How does the difference between rotations and sequences respond to model inputs ?
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}

# go through selected inputs and plot them on x with rot_minus_seq on y 
# initially with 200 reps doesn't show anything obvious, try it with 2000 overnight runs
in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')

@@

# PRCC

@@

#same as above
in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')

@@

# Runs where rotations are >20% better (in red) are associated with higher dominance of cost and lower dominance of selection

# but that does not guarantee, must be interaction with other inputs too.

@@

# looking at modelling it
# library(mgcv)
# mod_lm2 <- gam(rot_minus_seq ~ cost + dom_cos + dom_sel, data=df_res_all)
# summary(mod_lm2)  
# 
# mod_gam2 <- gam(rot_minus_seq ~ s(cost) + s(dom_cos) + s(dom_sel) + s(rr) + s(expo_hi) + s(male_expo_prop) + s(eff) + s(start_freqs), data=df_res_all)
# 
# summary(mod_gam2)  
# #Deviance explained = 19.5%
# 
# mod_gam3 <- gam(rot_minus_seq ~ te(cost,dom_cos,dom_sel), data=df_res_all)
# summary(mod_gam3)

#actually decision trees may be useful

@@

# Have a go at running for insecticides with different parameters
# starting to work ...
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=11, comment=NA}

# set random seed so that results can be reproduced
set.seed(2)
max_gen         <- 500
plot_scenarios  <- FALSE
n_scenarios     <- 100 #1000 #000

  #df_res_all <- data.frame()

  for(i in 1:n_scenarios)
  {
    #message should output to the R markdown console
    if (i%%200 == 1) message("scenario ",i," of ",n_scenarios," ",Sys.time())

    # inputs (could also be saved to a dataframe first & extracted one row at a time)
    linputs <- list()

    ## constant inputs ##
    linputs$coverage <- 1
    linputs$migration <- 0
    #linputs$cost <- 0 #variable cost below
    linputs$max_gen <- max_gen

    #beware min_rwr_interval, it's to stop repeated small rotations
    linputs$min_rwr_interval <- 5
    linputs$min_gens_switch_back <- 1
    linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
    linputs$no_r_below_mut <- TRUE #BEWARE trying out new mutation-selection balance option
    #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
    linputs$plot <-     FALSE

    ## variable inputs ##
    linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
    n_ins <- linputs$n_insecticides

    ## new those below can be different for each insecticide
    linputs$cost <-           runif(n_ins, min=0,   max=0.1) 
    #linputs$cost <-           0 
    linputs$expo_hi <-        runif(n_ins, min=0.4, max=0.9)  #min=0.1, max=0.9
    linputs$male_expo_prop <- runif(1, min=0,   max=1)
    linputs$eff <-            runif(n_ins, min=0.5, max=1) # min=0.3, max=1
    linputs$rr  <-            runif(n_ins, min=0.1, max=0.9)
    linputs$dom_sel <-        runif(n_ins, min=0,   max=1)
    linputs$dom_cos <-        runif(n_ins, min=0,   max=1) 
    #DIFFERENT
    linputs$start_freqs <-    runif(n_ins, min=0.0001,   max=0.01) #note not log yet

    # dispersal (initially with cost set to 0)
    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
    #linputs$migration <-      runif(1, min=0.1, max=0.9)

    # create null objects used later for plotting
    ggrot <- ggseq <- NULL

    # to compare sequence to rotation run all the scenarios above for both seq & rot 
    for(rot_or_not in 0:1)
    {  
      #to get rot_int of 0 every other run
      if (rot_or_not == 0) linputs$rot_interval <- 0
      else   #random rot interval integer
          linputs$rot_interval <-   sample( 5:50, 1 )  #beware set replace=T if more than 1

      ##################
      # run one scenario
      dfres <- do.call(run_rot, linputs)

      # summarise results per scenario
      # note this is also currently calc at end of run_rot() here is probably better

      res <- gens_under_thresh(dfres)

      # to plot results for checking
      if ( linputs$rot_interval == 0 )
      {
        gens_seq <- as.numeric(res)
      } else
      {
        gens_rot <- as.numeric(res)
      }

      if (plot_scenarios)
      {
        # make plot
        ggres <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=FALSE, plot=FALSE)

        # to plot results for checking
        if ( linputs$rot_interval == 0 )
        {
          ggseq <- ggres
          gens_seq <- as.numeric(res)
        } else
        {
          ggrot <- ggres
          gens_rot <- as.numeric(res)
        }
      }
    } #end rot_or_not loop

    # make an output row for each scenario with inputs, gens_rot, gens_seq
    # need to make sure I retain the rot_interval for rot

    # to cope with multiple values per scenario convert them to strings
    linputs2 <- lapply(linputs, function(x) ifelse(length(x)>1,toString(x),x))

    #added scenario id column
    df_in_out <- as.data.frame(c(linputs2, gens_rot=gens_rot, gens_seq=gens_seq, id=i), stringsAsFactors=FALSE)

    df_in_out$rot_minus_seq <- df_in_out$gens_rot - df_in_out$gens_seq

    #instead set up the dataframe to the required size in scenario1
    if (i==1)
    {
      df_res_all <- data.frame(matrix(NA, ncol=length(names(df_in_out)), nrow=n_scenarios))
      names(df_res_all) <- names(df_in_out)
    }

    #putting results into one row of overall results dataframe
    df_res_all[i,] <- df_in_out

    if (plot_scenarios)
    {  
      # plot input vals
      # extract out those between 0 & 1
      dfi <- data.frame(t(as.data.frame(linputs)))
      names(dfi) <- 'value'
      dfi$inputs <- rownames(dfi)

      ggins <- ggplot(dfi, aes(x=inputs,y=value)) + 
                geom_point() +
                theme_minimal() +
                ylim(0,1) +
                theme(axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1))

      grid.arrange(ggrot, ggseq, ggins, ncol=1)

      cat('scenario',i,'sum generations deployed insecticide under thresh rotation:', gens_rot,' sequence:',gens_seq,"\n\n")

      #can I give the code to rerun this scenario
      cat(paste0("run_rot(n_insecticides=",linputs$n_insecticides,", ",
                  "cost=",signif(linputs$cost,2),", ",
                  "start_freqs=",signif(linputs$start_freqs,2),", ", 
                  "rot_interval=",linputs$rot_interval,", ",
                  "eff=",signif(linputs$eff,2),",\n",
                  "dom_sel=",signif(linputs$dom_sel,2),", ",
                  "dom_cos=",signif(linputs$dom_cos,2),", ",
                  "rr=",signif(linputs$rr,2),", ",
                  "expo_hi=",signif(linputs$expo_hi,2),", ",
                  "coverage=",signif(linputs$coverage,2),",\n",
                  "migration=",signif(linputs$migration,2),", ",
                  "max_gen=",linputs$max_gen,", ", 
                  "min_rwr_interval=",linputs$min_rwr_interval,            
                  ")\n") )

    #save partial outputs just in case doesn't make to end
    if (i%%1000 == 0)
    {
      #df_res_all$rot_minus_seq <- df_res_all$gens_rot - df_res_all$gens_seq

      # save object containing inputs and results as rda for analysis
      save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))    
    } 
  } #end scenario loop

#to allow plotting of chosen scenarios by re-running them 
#initially didn't work for diff insecticides because the inputs are saved as strings 
#now the strings get converted back to vectors below

scenarios_to_plot <- c(21,23,32) #big difference ones all have insecticide1 as the good one
scenarios_to_plot <- c(29,38,46)
scenarios_to_plot <- c(13)

for(scenario in scenarios_to_plot)
{
  #cat(scenario,"\n")
  linputs <- as.list( df_res_all[scenario, 1:15] )

  #converting multiple insecticide inputs back from strings
  #have to ignore the ones that aren't strings

  #note the space after the comma in split=", "
  linputs <- lapply(linputs, function(x) if (!grepl(",",x)) x else
                                          as.numeric(unlist(strsplit(as.character(x),split=", "))))

  #TESTING changing min_rwr_interval from 5
  #if min_rwr_interval is set to 0 you get frequent switch backs to other insecticides for 1 generation
  #but when it's set to 5 it pushes frequences above the resistance freq threshold
  linputs$min_rwr_interval <- 5
  #min_gens_switch_back has to be 0 to allow one insecticide to continue when all others dead
  linputs$min_gens_switch_back = 0

  # run one scenario (rotation)
  dfres <- do.call(run_rot, linputs)

  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)

  # now run as sequence
  linputs$rot_interval <- 0
  dfres <- do.call(run_rot, linputs)

  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)

  grid.arrange(ggrot, ggseq, ncol=1)

>>>>

---
title: "sensi_an_rotations201911. To analyse rotation runs for paper"
output: pdf_document
#output: word_document
author: "Andy South, Sam Jones and Ian Hastings"
date: "`r Sys.Date()`"
---

Code to plot and analyse rotations runs for paper.

Runs themselves are done from rotation_runs2019.Rmd.

16 expts of 10,000 runs each comparing rotations and sequences.

Code below will need to be adapted to work on all 16 experiments, so far it just works on one.

@@

  # where the rotation runs saved
  infolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

@@

# load the data, start with just one experiment
# dataframes will need to be renamed because they will all be called df_res_all to start
infile <- "df_res_all_rot_mig1000.rda"
infile <- paste0(infolder, infile)

load(infile)

@@

@@

```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=3.5}
#eval false because smaller runs don't have these
#TODO change which runs to show as examples

@@

#Plots of how the difference between rotations and sequences respond to model inputs.

# go through selected inputs and plot them on x with rot_minus_seq on y 
# TODO check whether this covers all the inputs we want to look at
in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')

@@

# PRCC to show the sensitivity of the difference between rotation and sequence to inputs
# this can only be used for the experiments where all insecticides are the same
# (because in the ones where insecticides differ there are multiple values for each input)

@@

#same as above
#TODO check that this covers all the inputs we want to look at
in_cols <- c('cost','dom_cos','dom_sel','rr','expo_hi','male_expo_prop','eff','start_freqs')

@@

# Runs where rotations are >20% better (in red) are associated with higher dominance of cost and lower dominance of selection, but doesn't guarantee, must be interaction with other inputs too.
# this was for earlier runs when there was some strange model behaviour, may not be the case now
# and may not be very meaningful anyway

@@

# quick, not very successful look at modelling it
# library(mgcv)
# mod_lm2 <- gam(rot_minus_seq ~ cost + dom_cos + dom_sel, data=df_res_all)
# summary(mod_lm2)  
# 
# mod_gam2 <- gam(rot_minus_seq ~ s(cost) + s(dom_cos) + s(dom_sel) + s(rr) + s(expo_hi) + s(male_expo_prop) + s(eff) + s(start_freqs), data=df_res_all)
# 
# summary(mod_gam2)  
# #Deviance explained = 19.5%
# 
# mod_gam3 <- gam(rot_minus_seq ~ te(cost,dom_cos,dom_sel), data=df_res_all)
# summary(mod_gam3)

@@
